---
title: "spatial stuff"
author: "Rosemary Hartman"
date: "11/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(ggmap)
library(leaflet)
library(scales)
```

# Prepare regions shapefile

```{r}
ak_regions = read_sf("shapefile_demo_data/ak_regions_simp.shp")

plot(ak_regions)

st_crs(ak_regions)
class(ak_regions)

```

```{r}
ak_regions_338 = ak_regions %>% 
st_transform(crs = 3338)

plot(ak_regions_338)

#geomtry information carries no matter what operations you perform on it
str(ak_regions_338)
str(select(ak_regions_338, region))

```



# Prepare population data

```{r}
pop = read.csv("shapefile_demo_data/alaska_population.csv", stringsAsFactors = F)

head(pop)
class(pop)

```
Coerce our data frame to a sf file

```{r}
#we need to know the coordinate system. It's lat-long, so we know it
#is an unprojected coordinate system. IDeally it says it in the metadata
#but we'll just assume it is WGS1984
popsf = st_as_sf(pop, 
                 coords = c("lng", "lat"), #always x then y
                 crs = 4326, remove = F) 
#set remove = F to keep the origional lat long column
#you can also grab the coordinates out of the projection later.

st_crs(popsf)

```



# calculate population by region

```{r}
pop_joined = st_join(popsf, ak_regions_338, join = st_within)
#this doesn't work because we are in different coordinate reference systems

pop_3338 = popsf %>%
  st_transform(crs = 3338)
pop_joined = st_join(pop_3338, ak_regions_338, join = st_within)


```

```{r}
pop_region = pop_joined %>%
  as.data.frame() %>%
  group_by(region) %>%
  summarize(totalpop = sum(population))

pop_region

```

```{r}
pop_region_3338 = left_join(ak_regions_338, pop_region)
```

```{r}
pop_mgmt = pop_region_3338 %>%
  group_by(mgmt_area) %>%
  summarise(total_pop = sum(totalpop))

#the "summarize" function automatically takes the union of the features. 
#to stop this you can say "do_union = F"

plot(pop_mgmt["total_pop"])
```

```{r}
rivers_3338 = read_sf("shapefile_demo_data/ak_rivers_simp.shp")

st_crs(rivers_3338)

#if the EPSG code is not defined you can google the "proj4string" thing.

```



# make maps

```{r}
ggplot() +
  geom_sf(data =pop_region_3338, mapping =aes(fill = totalpop))+
  geom_sf(data = pop_3338, size = 0.5)+
  geom_sf(rivers_3338, mapping = aes(size = StrOrder), color = "black")+
  scale_size(range = c(0.01, 0.2), guide = F)+
  theme_bw() +
  scale_fill_continuous(name = "Total Population", low = "khaki", 
                        high = "firebrick", labels = comma)
```

Incorporate base maps using ggmap

We need to transform our data to 3857 - pseudomercator

```{r}
pop_3857 = pop_3338 %>%
  st_transform(crs = 3857)
```

A lot of web-based mapping systems are great for dirving directions but aren't great for accurate spatial computations

```{r}
# Define a function to fix the bbox to be in EPSG:3857
# See https://github.com/dkahle/ggmap/issues/160#issuecomment-397055208
ggmap_bbox_to_3857 <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))
  
  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))
  
  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}
```

```{r}
#upper left, lower right, 
bbox = c(-170, 52, -130, 64)
ak_map = get_stamenmap(bbox, zoom = 4)
ak_map_3857 = ggmap_bbox_to_3857(ak_map)

```

```{r}
ggmap(ak_map_3857) +
  geom_sf(data = pop_3857, aes(color = population), inherit.aes = F)
```

# Now make maps in leaflet

leaflet automatically projects it for you

```{r}
epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))
```

You need to give leaflet a crs and you need to give leaflet your data in 4326

```{r}
pop_region_4326 = pop_region_3338 %>%
  st_transform(crs = 4326)

pop_4326 = pop_3338 %>%
  st_transform(crs = 4326)
```

```{r}
leaflet(options = leafletOptions(crs = epsg3338)) %>%
  addPolygons(data = pop_region_4326, fillColor = "gray", weight = 1)
```


```{r}
pal <- colorNumeric(palette = "Reds", domain = pop_region_4326$totalpop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        addPolygons(data = pop_region_4326, 
                    fillColor = ~pal(totalpop),
                    weight = 1,
                    color = "black",
                    fillOpacity = 1,
                    label = ~region) %>% 
    addCircleMarkers(data = pop_4326,
                         lat = ~lat,
                         lng = ~lng,
                         radius = ~log(population/500), # arbitrary scaling
                         fillColor = "gray",
                         fillOpacity = 1,
                         weight = 0.25,
                         color = "black",
                         label = ~paste0(pop_4326$city, ", population ", comma(pop_4326$population))) %>%
        addLegend(position = "bottomleft",
                  pal = pal,
                  values = range(pop_region_4326$totalpop),
                  title = "Total Population")

m
```

