---
title: "Data tidying"
author: "Rosemary Hartman"
date: "11/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, message = F}

library(dplyr)
library(tidyr)

```


# Read and clean data

```{r}
#read in the data
catch_origional = read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors = F)

#check out the data
head(catch_origional)

```
Remove "all" and "notes" colomnes because i don't need them
```{r}
catch_data = catch_origional %>% 
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
select(-All, -notesRegCode)


head(catch_data)
```

```{r}
summary(catch_data)
```

## Fix Chinook Column

Use "mutate" to fix Chinook column

```{r}
catch_clean = catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))

summary(catch_clean)
```

Find the row that turned into NA

```{r}
i = which(is.na(catch_clean$Chinook))
i

catch_origional[401,]

```

```{r}
catch_clean = catch_data %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>%
  mutate(Chinook = as.numeric(Chinook))
```


# Reshape data

```{r}
catch_long = catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "species", values_to = "count")

head(catch_long)
```

Now I"ll move it back to wide format
```{r}
catch_wide = catch_long %>% 
  pivot_wider(names_from = Year, values_from = count)

head(catch_wide)

```

Data are actually in thousands of fish, so I'll rename the variable

```{r}
catch_long = catch_long %>% 
  rename(catch_thousands = count) %>% 
  mutate(catch = catch_thousands*1000) %>% 
  select(-catch_thousands)

head(catch_long)

```
# SUmmarize table.

```{r}
mean_region = catch_long %>% 
  group_by(Region, species) %>% 
  summarize(catch_mean = mean(catch), num_obs = n())

mean_region

total_sp = catch_long %>% 
  group_by(species) %>% 
  summarize(catch_sum = sum(catch, na.rm = T), num_obs = n()) %>% 
  arrange(desc(catch_sum))

total_sp
```
```

look at older data 
```{r}
total_spearly = catch_long %>% 
filter(Year <1900) %>%
  group_by(species) %>% 
  summarize(catch_sum = sum(catch, na.rm = T), num_obs = n()) %>% 
  arrange(desc(catch_sum))

total_spearly
```


# Joins


```{r}
region_defs = read.csv(url('https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1', method = "libcurl"), stringsAsFactors = F) %>%
select(code, mgmtArea) %>%
rename(Region = code)

head(region_defs)

```

```{r}
catch_joined = left_join(catch_long, region_defs)

head(catch_joined)
```



# Sepaate and unite
```{r}



site_codes = data.frame(site = c("HAW-100",
"HAW-101", "OAH-102", "OAH-103", "MAI-100"), stringsAsFactors = F)

site_codes_split = site_codes %>%
separate(site, into = c("island", "site_num"), sep = "-", remove = F)

head(site_codes_split)

```

```{r}
united = site_codes_split %>% 
unite(unitedcode, island, site_num, remove = F)

head(united)

```

